\chapter{Actionable semigroups}

Each definition containing the word ``semicategory'' will also implicity create a similar definition for semigroups.

\section{Category of ordered semicategory actions}

\begin{defn}
The category~$\mathbf{SemiCat}$ of \emph{semicategories} is small semicategories with semifunctors as morphisms.
\end{defn}

\begin{defn}
The category~$\mathbf{SemiCatAct}$ of \emph{semicategory actions} is small semicategories~$\mathcal{C}$ together with a semifunctor $\supfun{}:\mathcal{C}\to\mathbf{Set}$ with semifunctors as morphisms.
\end{defn}

\begin{defn}
The category~$\mathbf{OrdSemiCat}$ of \emph{ordered semicategories} is small ordered semicategories with monotone semifunctors as morphisms.
\end{defn}

\begin{defn}
The category~$\mathbf{SemiOrdSemiCatAct}$ of \emph{semi-ordered semicategory actions} is:
\begin{description}
\item[\textbf{objects}] small ordered semicategories~$\mathcal{C}$ together with a semifunctor $\supfun{}:\mathcal{C}\to\mathbf{Pos}$ conforming to the identity $a\sqsubseteq b\Rightarrow\supfun{a}x\sqsubseteq\supfun{b}x$;
\item[\textbf{morphisms}] monotone semifunctors between them.
\end{description}

In other words, $\mathbf{OrdSemiCatAct}$ is small semi-ordered semicategories together with a function~$\supfun{}$ to $\mathbf{Pos}$ with morphisms conforming to the identities:
\begin{enumerate}
\item $\supfun{b\circ a}=\supfun{b}\circ\supfun{a}$;
\item $a\sqsubseteq b\Rightarrow\supfun{a}x\sqsubseteq \supfun{b}x$.
\end{enumerate}

I call objects of this category \emph{semi-ordered semicategory actions}.
\end{defn}

\begin{defn}
\[ \mathbf{OrdSemiCatAct}\text{ is the wide subcategory of }\mathbf{SemiOrdSemiCatAct} \]
determined by the additional axiom $x\sqsubseteq y\Rightarrow\supfun{a}x\sqsubseteq\supfun{a}y$.

I call objects of this category \emph{ordered semicategory actions}.
\end{defn}

\begin{rem}
Contrary to sound mind, ordered semicategory actions (and ordered semigroup actions) were discovered only in 2019 (by me).
\end{rem}

\begin{defn}
I call objects of ordered semicategory actions \emph{interspaces} and elements of ordered semigroup actions \emph{spaces}.
\end{defn}

\begin{rem}
The usage of the word ``space'' in the above definition will be justified later.
\end{rem}

\section{Restricted identity transformations}

Fix an ordered semicategory action.

\begin{defn}
\emph{Semicategory action with identities} is an ordered semicategory~$S$ action~$\supfun{}$ together with a function $p\mapsto\id_p\in\End(A)$ for $p\in A$ for $A\in\Ob S$ such that whenever this equality is defined: \[ \supfun{\id_p}x=x\sqcap p. \]
\end{defn}

\begin{rem}
The following axioms often hold, too, but are not needed below:
\begin{enumerate}
\item $\id_p\circ f\sqsubseteq f$;
\item $f\circ\id_p\sqsubseteq f$;
\item $\id_p\circ\id_q=\id_{p\sqcap q}$.
\end{enumerate}
\end{rem}

\begin{prop}
\label{prop:id-inj}
$x\mapsto\id_x$ is an injection.
\end{prop}

\begin{proof}
Let $x\ne y$. Then $x\ne x\sqcap y$ (possibly, $x\sqcap y$ is undefined).
\[ \supfun{\id_y}x=x\sqcap y\ne x=x\sqcap x=\supfun{\id_x}x. \]
\end{proof}

\begin{prop}
For a $\sqcap$-se\-mi\-lat\-ti\-ce semicategory action, the semilattice axioms can be rewritten as:
\begin{enumerate}
\item $\supfun{\id_a}\supfun{\id_b}c = \supfun{\id_{\supfun{\id_a b}}}c$;
\item $\supfun{\id_a}b = \supfun{\id_b}a$;
\item $\supfun{\id_a}a = a$.
\end{enumerate}
\end{prop}

\emph{Restriction} of an interspace~$a$ to element~$x$ is $a|_x=a\circ\id_x$.

\emph{Square restriction} (a generalization of restriction of a topological space, metric space, etc.) of a space~$a$ to element~$x$ is $\id_x\circ a\circ\id_x$.

\section{Actionable semigroups}

\begin{defn}
An \emph{actionable semicategory} is a semicategory, and $\cdot$ is a binary operation on $A$ such that $(g\circ f)\cdot x = g\cdot(f\cdot x)$.
\end{defn}

\begin{defn}
An \emph{ordered actionable semicategory} is an actionable semicategory which is also an ordered semicategory with additional equation:
\begin{enumerate}
\item $f_0\sqsubseteq f_0\land g_1\sqsubseteq g_0 \Rightarrow g_0\cdot f_0\sqsubseteq g_1\cdot f_1$.
\end{enumerate}
\end{defn}

\begin{defn}
An \emph{actionable semicategory with identities} is an actionable semicategory that is also a semicategory with identities.
\end{defn}

\begin{rem}
An actionable semigroup is an universal algebra.
\end{rem}

\begin{rem}
Actionable ordered semigroups and actionable semigroups with identities are relational algebras.
\end{rem}

\begin{rem}
It is even an universal algebra signature\footnote{See here for \href{https://chatgpt.com/s/t_691e8f88da64819184591f9e310e1728}{properties of categories isomorphic to universal algebras}.} in the case of up or down semilattice.
\end{rem}

Let write the full universal algebra signature for actionable semigroup with identities ?? down-se\-mi\-lat\-ti\-ce:

\fxwarning{Move the axioms below.}
Operations $\circ$, $\cdot$, $p\mapsto\id_p$ with axioms:
\begin{enumerate}
\item $h\circ(g\circ f) = (h\circ g)\circ f$ (semigroup);
\item $h\cdot(g\cdot f) = (h\circ g)\cdot f$ (actionable);
\item $\id_a\cdot(\id_b\cdot c) = \id_{\id_a\cdot b}\cdot c$ ($\sqcap$-semilattice)\footnote{$a\sqcap(b\sqcap c) = (a\sqcap b)\sqcap c$};
\item $\id_a\cdot b = \id_b\cdot a$ (ditto)\footnote{$a\sqcap b = b\sqcap a$};
\item $\id_a\cdot a = a$ (ditto)\footnote{$a\sqcap a=a$};
\item $\id_{g_0\circ(\id_{\id_{f_0}\cdot f_1})}\cdot((\id_{g_0}\cdot g_1)\circ(\id_{f_0}\cdot f_1)) = (\id_{g_0}\cdot g_1)\circ(\id_{f_0}\cdot f_1)$ (ordered semigroup)\footnote{$(g_0\circ(f_0\sqcap f_1))\sqcap((g_0\sqcap g_1)\circ(f_0\sqcap f_1)) = (g_0\sqcap g_1)\circ(f_0\sqcap f_1)$};
\item $\id_{((\id_{f_1}\cdot f_0)\cdot g_0)}\circ((\id_{f_0}\cdot f_1)\cdot(\id_{g_0}\cdot g_1)) = (\id_{f_0}\cdot f_1)\circ(\id_{g_0}\cdot g_1)$ (ditto)\footnote{$((f_0\sqcap f_1)\circ g_0)\sqcap((f_0\sqcap f_1)\circ(g_0\sqcap g_1)) = (f_0\sqcap f_1)\circ(g_0\sqcap g_1)$};
\item $\id_{g_0\cdot(\id_{\id_{f_0}\cdot f_1})}\cdot((\id_{g_0}\cdot g_1)\cdot(\id_{f_0}\cdot f_1)) = (\id_{g_0}\cdot g_1)\cdot(\id_{f_0}\cdot f_1)$ (ordered actionable semigroup)\footnote{$(g_0\cdot(f_0\sqcap f_1))\sqcap((g_0\sqcap g_1)\cdot(f_0\sqcap f_1)) = (g_0\sqcap g_1)\cdot(f_0\sqcap f_1)$};
\item $\id_{((\id_{f_1}\cdot f_0)\cdot g_0)}\cdot((\id_{f_0}\cdot f_1)\cdot(\id_{g_0}\cdot g_1)) = (\id_{f_0}\cdot f_1)\cdot(\id_{g_0}\cdot g_1)$ (ditto)\footnote{$((f_0\sqcap f_1)\cdot g_0)\sqcap((f_0\sqcap f_1)\cdot(g_0\sqcap g_1)) = (f_0\sqcap f_1)\cdot(g_0\sqcap g_1)$}.
\end{enumerate}

For up-se\-mi\-lat\-ti\-ce, we need more axioms because it requires both $\sqcup$ and $\sqcap$ (for axiom~\ref{ax:ids}).

\section{From actionable semicategory to semicategory action}

\begin{defn}
If $(\dots,\circ,\cdot)$ is an actionable semicategory, then we can define the operation $\supfun{}$ by $\supfun{f}x = f\cdot x$.
\end{defn}

\begin{obvious}
The above defines an semicategory action with identities with operations~$\circ$ and~$\supfun{}$.
\end{obvious}

\begin{rem}
Semicategory actions defined this way have the sets of objects and morphisms equal to each other.
\end{rem}

\section{From semicategory action with identities to actionable semicategory}

Let $(A,\circ,\supfun{})$ be an ordered semicategory action with identities, each object set having greatest element~$\top$.

\begin{obvious}
$\supfun{f}x = \supfun{f}\supfun{\id_x}\top = \supfun{f\circ\id_x}\top = \im(f\circ\id_x)$.
\end{obvious}

By analogy with the above obvious, we can define $g\cdot f = \id_{\im(g\circ f)}$.

\begin{prop}
The above with the same semigroup operation~$\circ$ defines an actionable semicategory.
\end{prop}

\begin{proof}
Really, $(g\circ f)\cdot x = \id_{\im(g\circ f\circ x)} = \id_{\supfun{g}\supfun{f}\supfun{x}\top} = \id_{\supfun{g}{\im(f\circ x)}} = \id_{\supfun{g}(\top\sqcap\im(f\circ x))} = \id_{\im(g\circ(y\mapsto y\sqcap\im(f\circ x))} = \id_{\im(g\circ\id_{\im(f\circ x)})} = g\cdot(f\cdot x)$.

Monotonicity is obvious.
\end{proof}

\begin{rem}
Identities can't be the same, because the set of objects is not the same.
\end{rem}

\section{An injection}

Then \[ g\cdot\id_x = \supfun{g}\id_x = \id_{\im(g\circ\id_x)} = \id_{\supfun{g}x}. \]

So, we extended the mapping $\supfun{g}$ to a wider set, because $x\mapsto\id_x$ is an injection (see proposition~\ref{prop:id-inj}).

We can also restrict it to restricted identities: $\supfun{f}\id_x = f\cdot\id_x$ and obtain this way an ordered semicategory action isomorphic to the original one.

\fxnote{Is it a natural isomorphism?}

\section{Compositions of our ``inducing'' operations}

Every semicategory action with identities, whose objects have greatest elements, induces an actionable semicategory with identities, which induces another semicategory action with identities. Thus every semigroup with identities, whose objects have greatest elements, induces another semigroup with identities. The formula for this is $\supfun{g}'f = \id_{\im(g\circ f)}$, where $\supfun{}'$ is the induced action.

Let apply this formula two times:
\[ \supfun{g}''f = \id_{\im'(g\circ f)} = \id_{\supfun{g\circ f}'\top} = \id_{\im(g\circ f\circ\top)} = \id_{\im(g\circ f)} = \supfun{g}'f. \]

\begin{thm}
If we restrict the~$\cdot$ operation to the set that $\setcond{\id_x}{x\in A,A\in\Ob}$, then we recover the original semicategory action \fxnote{with identities?} up to \fxnote{Natural?} isomorphism of ??.
\end{thm}

\begin{proof}
??
\end{proof}

\section{Significance of ordered semicategory actions}

My research~\cite{volume-1} shows that all kinds of spaces met in general topology (topological spaces, metric spaces, uniform spaces, locales and frames, etc.) are kinds of ordered semicategory actions.

However, because that is a draft, I am still ``confused'' about what kind of category embeddings they are: category isomorphism, category equivalence, or like this.

I will publish the details later.

In any case, that the entire theory of general topology is reducible to ordered semicategory actions, makes ordered semicategory actions immensely important, about on-par with group theory.

As I showed above that ordered semigroup actions (when identities are defined) are reducible to universal algebra, this makes all general topology reducible to universal algebra, what also seems to be very important.

It is yet unclear for me, whether non-ge\-ne\-ral-to\-po\-lo\-gy spaces such as Euclidean spaces are also reducible to ordered semigroup actions, so possibly defining ``space in general''.